---
title: "Animals Adoptability Analysis"
author: "Xiao Li"
output: 
  flexdashboard::flex_dashboard:
  vertical_layout: scroll
source_code: embed
---

```{r , include=FALSE}
library(readr)
library(tidyverse)
library(gapminder) 
library(dplyr)
library(ggplot2)
library(broom)
library(glmnet)
library(caret)
library(ISLR)
library(janitor)
library(rpart)
library(rpart.plot)
library(partykit)
library(nnet)
library(MASS)
library(wordcloud)
library(RColorBrewer)
library(randomForest)
library(gbm)
library(plotROC)
library(kernlab)
library(nnet)
library(stringi)
library(tm)
library(gridExtra)
theme_set(theme_bw())
```  

```{r setup, include=FALSE}
# Data
df <- read_csv("train.csv")

df <- df %>%
  filter(Quantity == "1")%>%
  subset(select = -c(Name, PetID, Quantity, State, RescuerID))
```

```{r word cloud, include=FALSE}

Description <- stringi::stri_trans_general(df$Description, "latin-ascii")

description<-Corpus(VectorSource(Description))

description <- tm_map(description,stripWhitespace)
description <- tm_map(description,tolower)
description <- tm_map(description,removeNumbers)
description <- tm_map(description,removePunctuation)
description <- tm_map(description,removeWords, stopwords("english"))
description <- tm_map(description, removeWords, 
  c("and","the","our","that","for","are","also","more","has","must","have","should","this","with","dog","cat","dogs","cats","home","adoption","can","give"))
```

```{r, include=FALSE}
# Data without `Description`.
df1 <- subset(df, select = -c(Description))

# Data for tree
df2 <- df1 %>%
  mutate(AdoptionSpeed = as.factor(AdoptionSpeed))
```
Row {.tabset .tabset-fade}
-------------------------------------

### Word Cloud: Description of Animals' Profiles
  
```{r}
wordcloud (description, scale=c(3.5,0.5), max.words=100, random.order=FALSE, rot.per=0.35, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
```

### Adoption Speed

```{r}
p <- ggplot(df1, aes(x = AdoptionSpeed)) + geom_bar(fill = "dodgerblue4")
p
```

### Age Impact

```{r}
p1 <- ggplot(df1, aes(x = Age, y = AdoptionSpeed)) + geom_point(alpha = .15, col = "#6e0000") + 
  ggtitle("Age vs. Adoption Speed")
p2 <- ggplot(df1, aes(x = Age)) + geom_histogram(fill = "dodgerblue4") + ggtitle("Age Distribution")

grid.arrange(p1, p2, nrow = 1)
```

### Video Amt / Photo Amt Impact
  
```{r}
p1 <- ggplot(df1, aes(x = VideoAmt, y = AdoptionSpeed)) + 
  geom_point(alpha = .15, col = "#6e0000") + ggtitle("Video Amt vs. Adoption Speed") 

p2 <- ggplot(df1, aes(x = PhotoAmt, y = AdoptionSpeed)) + 
  geom_point(alpha = .15, col = "#6e0000") + ggtitle("Photo Amt vs. Adoption Speed")

grid.arrange(p1, p2, nrow = 1)
```


### Breed Impact

```{r}
ggplot(df1, aes(x = Breed1, y = AdoptionSpeed)) + geom_point(alpha = .15, col = "#6e0000") +
  labs(title="Breed vs. Adoption Speed")
```

### Maturity Size Impact

```{r}
p1 <- ggplot(df1, aes(x = MaturitySize, y = AdoptionSpeed)) + geom_point(alpha = .15, col = "#6e0000")+
  ggtitle("Maturity Size vs. Adoption Speed")
p2 <- ggplot(df1, aes(x = MaturitySize)) + geom_bar(fill = "dodgerblue4") + labs(title="Maturity Size")

grid.arrange(p1, p2, nrow = 1)
```

### Healthy Condition Impact  

```{r}
p1 <- ggplot(df1, aes(x = Health, y = AdoptionSpeed)) + geom_point(alpha = .15, col = "#6e0000") +
  ggtitle("Health vs. Adoption Speed")
p2 <- ggplot(df1, aes(x = Health)) + geom_bar(fill = "dodgerblue4") + labs(title = "Healthy Condition")

grid.arrange(p1, p2, nrow = 1)
```

### Other Characteristics: Animal Characteristics

```{r}
Type <- ggplot(df1, aes(x = Type)) + geom_bar(fill = "dodgerblue4") + ggtitle("Type")
Gender <- ggplot(df1, aes(x = Gender)) + geom_bar(fill = "dodgerblue4") + labs(title = "Gender")
FurLength <- ggplot(df1, aes(x = FurLength)) + geom_bar(fill = "dodgerblue4") + ggtitle("Fur Length")

grid.arrange(grobs=list(Type, Gender, FurLength), nrow(1))
```

### Other Characteristics: Health Conditions
```{r}
Vaccinated <- ggplot(df1, aes(x = Vaccinated)) + geom_bar(fill = "dodgerblue4") + ggtitle("Vaccinated")
Dewormed <- ggplot(df1, aes(x = Dewormed)) + geom_bar(fill = "dodgerblue4") + ggtitle("Dewormed")
Sterilized <- ggplot(df1, aes(x = Sterilized)) + geom_bar(fill = "dodgerblue4") + ggtitle("Sterilized")

grid.arrange(grobs = list(Vaccinated, Dewormed, Sterilized), nrow(1))
```


### Classification Tree: Best Tree  
```{r}
set.seed(2)
inTraining <- createDataPartition(df2$AdoptionSpeed, p = .7, list = F)
training <- df2[inTraining, ]
testing  <- df2[-inTraining, ]
fit_control <- trainControl(method = "repeatedcv",
                            number = 10, 
                            repeats = 10)
cv_adop_tree <- train(AdoptionSpeed ~ ., 
                     data = training,
                     method = "rpart", 
                     trControl = fit_control)
cv_adop_tree
```

### Classification Tree: Tree Plot
```{r}
plot(as.party(cv_adop_tree$finalModel))
```

### Bagging
```{r}
set.seed(10982)
adop_bag <- randomForest(AdoptionSpeed ~ ., data = training, mtry = 17) 
adop_bag         
```


### Best Method: Random Forest w/ mtry = 2
```{r}
rf_adop_2 <- randomForest(AdoptionSpeed ~.,
                          data = training,
                          mtry = 2)
rf_adop_2
```

